### 2019-03-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Should reaaaaaallllyyyy go

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
On the first hand,
the \CUSUM{} test is rather complicated and parametric, as it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test is similarly complex, and it is also studied in \cite{LiuLeeShroff17},
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments,
and even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
On the other hand, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the different of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).

Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
%
We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\lceil 1/\alpha \rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
%
A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$. Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{align*}
 &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
%
\begin{eqnarray*}
R_T \leq T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} + \ \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{eqnarray*}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

@@

\begin{assumption}\label{ass:6:LongPeriodsGlobal}
    Define
    \begin{equation}\label{def:6:DelayGlobal}
        d^{(k)} = d^{(k)}(\alpha,\delta) = \lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \rceil.
    \end{equation}
    %
    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
\end{assumption}

@@

We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
%
\begin{eqnarray}\label{eq:6:GeneRegretBound}
    R_T &\leq& T \bP\left(\cE_T^c\right) + \alpha T + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
    & & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{eqnarray}

@@

Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}. A first key observation is
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:

\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.

>>>>

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
On the first hand,
the \CUSUM{} test is rather complicated and parametric, as it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test is similarly complex, and it is also studied in \cite{LiuLeeShroff17},
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments,
and even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
On the other hand, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the different of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

The forced exploration scheme used in \GLRklUCB{} (lines $3$-$5$) generalizes the deterministic exploration scheme proposed for \MUCB{} by (\cite{CaoZhenKvetonXie18}), whereas \CUSUMUCB{} performs randomized exploration.
%
We mimic the deterministic exploration introduced for \MUCB, while \CUSUMUCB{} is using a uniform exploration with probability $\alpha$.
Both approaches also require a mechanism to add exploration, parameterized by $\alpha\in(0,1)$.
\CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.
%
A consequence of this forced exploration is given in Proposition~\ref{prop:6:EnoughSamples}, proved below.

@@

Recall that the regret is defined as $R_T = \bE\left[\sum_{t=1}^T(\mu_{i^*_t}(t) - \mu_{I_t}(t))\right]$. Introducing $\cD(T,\alpha)$ the (deterministic) set of time steps at which the forced exploration is performed before time $T$ (see lines $3$-$4$ in Algorithm~\ref{algo:6:GLRklUCB}), one can write, using notably that $\mu_{i^*_t} - \mu_{I_t} \leq 1$, due to the bounded rewards,
\begin{align*}
 &\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))  \leq  \sum_{t=1}^{T} \indic(t \in \cD(T,\alpha)) + \sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(t \notin\cD(T,\alpha), \UCB_{I_t}(t) \geq \UCB_{i^*_t}(t)\right) \\
 & \hspace{1cm}\leq  \alpha T + \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right) + \sum_{i=1}^K\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{i}(t))\indic\left(I_t = i, \UCB_{i}(t) \geq \mu_{i^*_t}(t)\right).
\end{align*}
%
Introducing some \emph{good event} $\cE_T$ to be specified in each case, one can write
%
\begin{align*}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T & + \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{align*}
%
Each analysis requires to define an \emph{appropriate good event}, stating that \emph{some} change-points are detected within a reasonable delay. Each regret bound then follows from upper bounds on term $(A)$, term $(B)$, and on the failure probability $\bP(\cE_T^c)$.
%
To control $(A)$ and $(B)$, we split the sum over consecutive segments $[\tau^{(k)}+1, \tau^{(k+1)}]$ for global changes and $[\tau^{(k)}_i+1, \tau^{(k+1)}_i]$ for each arm $i$ for local changes, and use elements from the analysis of \klUCB{} of \cite{KLUCBJournal}.

@@

\begin{assumption}\label{ass:6:LongPeriodsGlobal}
    Define
    \begin{equation}\label{def:6:DelayGlobal}
        d^{(k)} = d^{(k)}(\alpha,\delta) = \left\lceil \frac{4K}{\alpha\left(\Delta^{(k)}\right)^2}\beta(T,\delta) + \frac{K}{\alpha} \right\rceil.
    \end{equation}
    %
    Then we assume that for all $k \in \{1,\dots,\Upsilon_T\}$,
    $\tau^{(k)} - \tau^{(k-1)} \geq 2\max (d^{(k)},d^{(k-1)})$.
\end{assumption}

@@

We gave in Section~\ref{sec:6:RegretAnalysis} the following decomposition of the regret $R_T$,
%
\begin{align}\label{eq:6:GeneRegretBound}
    R_T \leq T \bP\left(\cE_T^c\right) + \alpha T &+ \underbrace{\bE\left[\indic(\cE_T) \sum_{t=1}^T \indic\left(\UCB_{i^*_t}(t) \leq \mu_{i^*_t}(t)\right)\right]}_{(A)} \nonumber \\
    & + \underbrace{\bE\left[\indic(\cE_T)\sum_{t=1}^T (\mu_{i^*_t}(t) - \mu_{I_t}(t))\indic\left(\UCB_{I_t}(t) \geq \mu_{i^*_t}(t)\right)\right]}_{(B)}.
\end{align}

@@

Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}. A first key observation is
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
